### 2020-02-27 Writing/disparity_review.tex
Updated Natalie's comments

\noindent Clades of organisms are characterised by variation in both numbers of species and range of phenotypes through time.
At the extremes, clades may be exceptionally species-rich and phenotypically diverse (e.g. cichlids or molluscs), species-rich but phenotypically conservative (e.g. rodents or nematodes), species-poor but phenotypically diverse (e.g. Afrotheria), % NC: Can we expand Afrotheria a bit? All the other examples are using common names. Can we use common names of afrotheres? Or at least say afrotherian mammals?
or depauperate in both species and phenotypic diversity (e.g. lungfish).
These phenomena suggest that taxonomic and phenotypic diversity are not necessarily linked, raising important questions about how phenotypic diversity arises, such as:
How does diversity evolve?
Are some morphologies more common than others?
Can anatomy evolve in all ``directions'' or are some anatomies impossible to achieve? (Mike Foote 1997, 1995) % NC: Seems weird only having these two refs here and no refs in the other questions. Remove?
What role does ecology play in structuring morphological diversity?
Analyses of species diversity have a venerable history, but those of phenotypic diversity (hereafter \emph{disparity}) are comparatively more recent. 
Originally defined as ``multidimensional morphological dissimilarity at a macroevolutionary scale'' \citep{runnegar1987rates,Gould1991-nh} the concept of disparity emerged from attempts by palaeobiologists to characterise the evolutionary origin of animal bodyplans and from attempts by comparative developmental biologists to provide causal explanations for their emergence.
However, analyses of ``disparity'' have since expanded into comparative biology as a means of capturing the effect of intrinsic and extrinsic causal agents in morphological evolution.
Typically, methods to capture disparity are based on multidimensional spaces where each dimension represents an aspect of morphological variation (a trait) and biological observations (e.g. taxa) can be placed in this space based on their trait values.
Such multidimensional spaces (or morphospaces) can then be used to tackle a diverse array of questions that can be grouped into four main (non-mutually exclusive) classes (Fig. 1):

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/figure_disparities.pdf}
\caption{
    The four main types of disparity analysis. \textbf{Descriptive disparity} focuses on describing the features of morphospace occupation; \textbf{disparity-through-time} investigates the evolution of the morphospace through time including the effect of extinction events; \textbf{disparity and taxonomic diversity} compares different measures of biodiversity; \textbf{disparity as a proxy for ecology} uses disparity as a proxy for the ecological or functional role of a group.
    These categories are not independent and many disparity studies will cover more than one.
    %@@@ Add disparity and evo-devo
    % NC: I removed the mention of constraints because I hate the idea it's all constraints - there are other things at play why are we only mentioning constraints? (I suspect Anjali or Vera put that in?) Also it didn't fit with the other examples which don't mention mechanisms...
}
\label{Fig:disparity}
\end{figure}

>>>>

\noindent Clades of organisms are characterised by variation in both numbers of species and range of phenotypes through time.
At the extremes, clades may be exceptionally species-rich and phenotypically diverse (e.g. cichlids or molluscs), species-rich but phenotypically conservative (e.g. rodents or nematodes), species-poor but phenotypically diverse (e.g. afrotherian mammals), or depauperate in both species and phenotypic diversity (e.g. lungfish).
These phenomena suggest that taxonomic and phenotypic diversity are not necessarily linked, raising important questions about how phenotypic diversity arises, such as:
How does diversity evolve?
Are some morphologies more common than others?
Can anatomy evolve in all ``directions'' or are some anatomies impossible to achieve?
What role does ecology play in structuring morphological diversity?
Analyses of species diversity have a venerable history, but those of phenotypic diversity (hereafter \emph{disparity}) are comparatively more recent. 
Originally defined as ``multidimensional morphological dissimilarity at a macroevolutionary scale'' \citep{runnegar1987rates,Gould1991-nh} the concept of disparity emerged from attempts by palaeobiologists to characterise the evolutionary origin of animal bodyplans and from attempts by comparative developmental biologists to provide causal explanations for their emergence.
However, analyses of ``disparity'' have since expanded into comparative biology as a means of capturing the effect of intrinsic and extrinsic causal agents in morphological evolution.
Typically, methods to capture disparity are based on multidimensional spaces where each dimension represents an aspect of morphological variation (a trait) and biological observations (e.g. taxa) can be placed in this space based on their trait values.
Such multidimensional spaces (or morphospaces) can then be used to tackle a diverse array of questions that can be grouped into four main (non-mutually exclusive) classes (Fig. 1):

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/figure_disparities.pdf}
\caption{
    The four main types of disparity analysis. \textbf{Descriptive disparity} focuses on describing the features of morphospace occupation; \textbf{disparity-through-time} investigates the evolution of the morphospace through time including the effect of extinction events; \textbf{disparity and taxonomic diversity} compares different measures of biodiversity; \textbf{disparity as a proxy for ecology} uses disparity as a proxy for the ecological or functional role of a group.
    These categories are not independent and many disparity studies will cover more than one.
}
\label{Fig:disparity}
\end{figure}
